FHJ: A Formal Model for Hierarchical Dispatching and Overriding

### 2018-05-08 ECOOP18/sections/Discussion.tex
labeled. Ready for Bruno review

\section{Discussion}\label{sec:discussion}
In this section, we will talk about the prototype implementation and discuss the design space and reflect about some of the design decisions of our work. We relate our language to traits, Java interfaces as well as other languages. Furthermore, we discuss ways to improve our work.

\subsection{Implementation}
We have a simple implementation of \name{} interpreter in Scala. The implementation validates all the examples presented in the paper. The implementation is done in 
Scala, strictly following the type-checking rules and small-step semantics
presented in the paper. One nice feature of the implementation is that it can
show the detailed step-by-step evaluation of the program, which is convenient for
understanding and debugging programs \& semantics. The final model of \name{}
is actually a bit different from our initial models. This is partly because in
the process of implementing and proving theorems we found errors and modified
the earlier models, creating a positive feedback circle.

\subsection{Abstract Methods}

Abstract methods are one of the key features in most general OO languages. For example, Java interfaces were designed
to include only method declarations, and those abstract methods can be implemented in a class body. 
The formal Featherweight Java model~\cite{Igarashi01FJ} does not include abstract methods because of the orthogonality
to the core calculus. In traits, a
similar idea is to use keywords like ``\textbf{require}'' for abstract method declarations~\cite{scharli03traits}.
Abstract methods provide a way to
delay the implementations to future subtypes. Using overriding, they also help to ``exclude'' existing implementations.

@@

One non-orthogonal extension to \MIM{} could be to generalize the model to allow multiple hierarchical method overrding, meaning that, we allow overriding methods to update multiple branches instead of only one branch. It offers a more fine-grained mechanism for merging, and can be helpful to easily understand the structure of the hierarchy. Multiple overriding would be useful in the following situation, for example:

>>>>

\section{Discussion}\label{sec:discussion}
In this section, we will discuss the design space and reflect about some of the design decisions of our work. We relate our language to traits, Java interfaces as well as other languages. Furthermore, we discuss ways to improve our work.

\subsection{Abstract Methods}
Abstract methods are one of the key features in most general OO languages. For example, Java interfaces were designed
to include only method declarations, and those abstract methods can be implemented in a class body. 
The formal Featherweight Java model~\cite{Igarashi01FJ} does not include abstract methods because of the orthogonality
to the core calculus. In traits, a
similar idea is to use keywords like ``\textbf{require}'' for abstract method declarations~\cite{scharli03traits}.
Abstract methods provide a way to
delay the implementations to future subtypes. Using overriding, they also help to ``exclude'' existing implementations.

@@

\yanlin{Bruno: discussed multiple overriding and the two designs}
One non-orthogonal extension to \MIM{} could be to generalize the model to allow multiple hierarchical method overrding, meaning that, we allow overriding methods to update multiple branches instead of only one branch. It offers a more fine-grained mechanism for merging, and can be helpful to easily understand the structure of the hierarchy. Multiple overriding would be useful in the following situation, for example:
